---
title: "IMB Employee Attrition"
author: "yingrui Lyu"
date: "2020/12/20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
---
#reload needed package
```{r packages, message=FALSE, warning=FALSE}
library(ggplot2)
library(grid)
library(gridExtra)
library(plyr)
library(rpart)
library(rpart.plot)
library(randomForest)
library(caret)
library(gbm)
library(survival)
library(pROC)
library(DMwR)
library(scales)
```
---
#set a workpath to red the data
```{r}
Attr.df <- read.csv('D:/Employee_Attritiondata/EmployeeAttrition.csv')
head(Attr.df)
summary(Attr.df)
```
---
#3.Univariate Analysis
3.1 Explore the analysis of gender, age, length of service, number of companies, and time limit in the company
# g1 Employee’s age distribution
# g2 The relationship between the number of companies that the quit employee has worked
# g3 The distribution of quit employees’ service age
# g4 Distribution of the overall working years of employees leaving
```{r}
g1 <- ggplot(Attr.df,aes(x=Age,fill=Attrition))+
  geom_density(alpha=0.7)

g2 <- ggplot(Attr.df,aes(x=NumCompaniesWorked,fill=Attrition))+
  geom_density(alpha=0.7)

g3 <- ggplot(Attr.df,aes(x=YearsAtCompany,fill=Attrition))+
  geom_density(alpha=0.7)

g4 <- ggplot(Attr.df,aes(x=TotalWorkingYears,fill=Attrition))+
  geom_density(alpha=0.7)
grid.arrange(g1,g2,g3,g4,ncol=2,nrow=2)
```
---

#g5 Gender distribution of resigning employees
# g6 Position level distribution of resigning employees 
# g7 Educational background distribution of resigning employees 
# g8 Departmental distribution of resigning employees
```{r}
g5 <- ggplot(Attr.df, aes(x= Gender,fill = Attrition)) + 
  geom_bar(position = "fill") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 


g6 <-ggplot(Attr.df, aes(x= JobLevel,fill = Attrition)) + geom_bar(position = "fill") + labs(y="Percentage") + scale_y_continuous(labels=percent) 

g7 <- ggplot(Attr.df, aes(x= Education,fill = Attrition)) +geom_bar(position = "fill") + labs(y="Percentage") + scale_y_continuous(labels=percent) 
 
g8_ <- ggplot(Attr.df, aes(x= Department,fill = Attrition)) + geom_bar(position = "fill") + labs(y="Percentage") + scale_y_continuous(labels=percent) 
g8=g8_+ theme(axis.text.x = element_text(angle = 15,hjust=1))
grid.arrange(g5, g6, g7, g8, ncol = 2, nrow = 2)
```
---
# g11 The relationship between resigning employees and salary increase 
# g12 The relationship between resigning employees and the number of training 
# g13 The relationship between resigning employees and annual promotion 
# g14 The relationship between resigning employees and stock options
```{r}
g11 <- ggplot(Attr.df, aes(x = PercentSalaryHike, fill = Attrition)) + 
  geom_density(alpha = 0.7)

g12 <- ggplot(Attr.df, aes(x= TrainingTimesLastYear,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 

g13 <- ggplot(Attr.df, aes(x = YearsSinceLastPromotion, fill = Attrition)) + 
  geom_density(alpha = 0.7)

g14 <- ggplot(Attr.df, aes(x= StockOptionLevel,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 

grid.arrange(g11, g12, g13, g14, ncol = 2)
```
---
# g15 The relationship between resigning employees and job satisfaction 
# g16 The relationship between resigning employees and colleague satisfaction 
# g17 The relationship between resigning employees and job satisfaction
```{r}
g15 <- ggplot(Attr.df, aes(x= JobSatisfaction,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 

g16 <- ggplot(Attr.df, aes(x= RelationshipSatisfaction,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 

g17 <- ggplot(Attr.df, aes(x= EnvironmentSatisfaction,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 
grid.arrange(g15, g16,g17, ncol = 3)
```

---
# g18 The relationship between resigned employees and overtime 
# g19 The relationship between resigned employees and work life 
# g20 The relationship between resigned employees and business trip 
# g21 The relationship between resigned employees and working distance
```{r}
g18 <- ggplot(Attr.df, aes(x= OverTime,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 

g19 <- ggplot(Attr.df, aes(x= WorkLifeBalance,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 

g20_ <- ggplot(Attr.df, aes(x= BusinessTravel,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 
g20=g20_+ theme(axis.text.x = element_text(angle = 30, hjust=1))#transfer x.lable axis

g21 <- ggplot(Attr.df,aes(x=DistanceFromHome,fill=Attrition))+
  geom_density(alpha=0.7)

grid.arrange(g18, g19,g20,g21, ncol = 2)
```

---
# g9 The relationship between resigned employees and monthly salary 
# g10 The relationship between resigned employees and job level
```{r}
g9 <- ggplot(Attr.df,aes(x=MonthlyIncome,fill=Attrition))+
  geom_density(alpha=0.7)


g10 <- ggplot(Attr.df, aes(x= JobInvolvement,  group=Attrition)) + 
  geom_bar(aes(y = ..prop.., fill = Attrition), 
           stat="count", alpha = 0.7,position = "identity",color="black") +
  labs(y="Percentage") + scale_y_continuous(labels=percent) 

grid.arrange(g9, g10, ncol = 2)
```
---
#3.6 Further analysis of the relationship between monthly salary and job level
```{r}
ggplot(Attr.df,aes(x=JobInvolvement,y=MonthlyIncome,group=JobInvolvement))+
  geom_boxplot(aes(fill=factor(..x..)),alpha=0.7)+
  theme(legend.position = 'none',plot.title = element_text(hjust = 0.5))+
  facet_grid(~Attrition)+ggtitle('Attrition')
```
---
# Remove unnecessary reasons in the data set
```{r}
levels(Attr.df$JobRole) <- c("HC", "HR", "Lab", "Man", "MDir", "RsD", "RsSci", "SlEx", "SlRep")
levels(Attr.df$EducationField) <- c("HR", "LS", "MRK", "MED", "NA", "TD")
Attr.df <- Attr.df[c(-9,-10,-22,-27)]
```

# Divide the data set into training set and test set
```{r}
n <- nrow(Attr.df)
rnd <- sample(n,n*0.7)
train <- Attr.df[rnd,]
test <- Attr.df[-rnd,]
```
# build GBM model
```{r}
set.seed(3443)
ctrl <- trainControl(method = 'cv',number=10,summaryFunction = twoClassSummary,classProbs = T)
gbmfit <- train(Attrition~.,data=train,method='gbm',verbose=F,metric='ROC',trControl=ctrl)
gbmpreds <- predict(gbmfit,test)
Atr_factor<-factor(test$Attrition)#为了避免将字符串改为数字过程中产生NA，先用factor函数转换
gbmpreds_factor<-factor(gbmpreds)
rocgbm <- roc(as.numeric(Atr_factor),as.numeric(gbmpreds_factor), plot=TRUE, print.thres=TRUE, print.auc=TRUE)
rocgbm$auc
prop.table(table(test$Attrition, gbmpreds, dnn = c("Actual", "Predicted")),1)
```
---
#need to optimize GBM
# to set the same seed as fomer
```{r}
ctrl$seeds <- gbmfit$control$seeds
```
# weighted random sampling-GBM
```{r}
model_weights <- ifelse(train$Attrition == 'No',
                        (1/table(train$Attrition)[1]),
                        (1/table(train$Attrition)[2]))


weightedleft <- train(Attrition ~ .,
                      data=train,
                      method='gbm',
                      verbose=F,
                      weights=model_weights,
                      metric='ROC',
                      trControl=ctrl)

weightedpreds <- predict(weightedleft,test)
Atr_factor<-factor(test$Attrition)
weightedpreds_factor<-factor(weightedpreds)
rocweight <- roc(as.numeric(Atr_factor),as.numeric(weightedpreds_factor))
plot(rocweight)
rocweight$auc
```
# up_sampling
```{r}
ctrl$sampling <- 'up'
set.seed(3433)
upfit <- train(Attrition ~., 
               data = train, 
               method = "gbm", 
               verbose = FALSE, 
               metric = "ROC", 
               trControl = ctrl)

uppreds <- predict(upfit, test)
#Atr_factor<-factor(test$Attrition)
uppreds_factor<-factor(uppreds)
rocup <- roc(as.numeric(Atr_factor), as.numeric(uppreds_factor))
rocup$auc
prop.table(table(test$Attrition, uppreds, dnn = c("Actual", "Predicted")),1)
```
# down_sampling
```{r}
ctrl$sampling <- 'down'
set.seed(3433)
downfit <- train(Attrition ~., 
                 data = train, 
                 method = "gbm", 
                 verbose = FALSE, 
                 metric = "ROC", 
                 trControl = ctrl)

downpreds <- predict(downfit, test)
Atr_factor<-factor(test$Attrition)
downpreds_factor<-factor(downpreds)
rocdown <- roc(as.numeric(Atr_factor), as.numeric(downpreds_factor))
rocdown$auc
prop.table(table(test$Attrition, downpreds, dnn = c("Actual", "Predicted")),1)
prop.table(table(test$Attrition, weightedpreds, dnn = c("Actual", "Predicted")),1)
```
---

# Use model to predict turnover 
#5.1 See what factors affect employee turnover
```{r}
varImp(upfit)
```

##Predict the turnover rate of employees with high work input and low monthly salary
```{r}
upfitprobs <- predict(upfit,test,type = 'prob')
test$Prediction <- upfitprobs$Yes
ggplot(test,
       aes(x=MonthlyIncome,y=Prediction,color=factor(JobInvolvement)))+
  geom_point(alpha=0.7)+
  geom_smooth(method = 'lm')+
  facet_wrap(~JobInvolvement)+
  theme(legend.position = 'none')+
  ggtitle('JobInvolvement')+
  theme(plot.title = element_text(hjust = 0.5))
```

#Forecast those positions with the highest turnover rate
```{r}
p=ggplot(test,aes(x=JobRole,y=Prediction,fill=JobRole))+
  geom_boxplot(alpha=0.5)+
  theme(legend.position = 'none')+
  scale_y_continuous(labels = percent)
p+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))#transfer x.lable axis
```
